Palmer Archipelago Penguins Data in the palmerpenguins R Package - An Alternative to Anderson's Irises

### 2021-04-30 penguins.Rmd
change RJbibliography name to penguins.bib for consistency (#25)
reknit and repurl (#26)

---
title: Palmer Archipelago Penguins Data in the palmerpenguins R Package - An Alternative to Anderson's Irises
author:
  - name: Allison M. Horst
    affiliation: Bren School of Environmental Science and Management
    address:
    - University of California, Santa Barbara
    - Santa Barbara, CA 93106-5131
    email:  ahorst@ucsb.edu
  - name: Alison Presmanes Hill
    affiliation: RStudio, PBC
    address:
    - 250 Northern Ave
    - Boston, MA 02210
    email:  alison@rstudio.com
  - name: Kristen B. Gorman
    affiliation: College of Fisheries and Ocean Sciences
    address:
    - University of Alaska Fairbanks 
    - 2150 Koyukuk Drive
    - 245 O’Neill Building
    - Fairbanks, AK 99775-7220
    email:  kbgorman@alaska.edu
abstract: >

 In 1935, Edgar Anderson collected size measurements for 150 flowers from three species of *Iris* on the Gaspé Peninsula in Quebec, Canada. Since then, Anderson's *Iris* observations have become a classic dataset in statistics, machine learning, and data science teaching materials. It is included in the base R **datasets** package as **iris**, making it easy for users to access without knowing much about it. However, the lack of data documentation, presence of non-intuitive variables (e.g. "sepal width"), and perfectly balanced groups with zero missing values make **iris** an inadequate and stale dataset for teaching and learning modern data science skills. Users would benefit from working with a more representative, real-world environmental dataset with a clear link to current scientific research. Importantly, Anderson’s *Iris* dataset appeared in a 1936 publication by R. A. Fisher in the *Annals of Eugenics* (which is often the first-listed citation for the dataset), inextricably linking **iris** to racist research and theory. Thus, a modern alternative to the **iris** dataset is needed. In this paper, we introduce the **palmerpenguins** R package, which includes morphological measurements collected from 2007 - 2009 for three species of *Pygoscelis* penguins that breed on islands throughout the Palmer Archipelago, Antarctica. The **penguins** data provides an approachable, charismatic, and near drop-in replacement for **iris** with topical relevance global climate change and environmental impacts on marine predators. Since the release on CRAN in July 2020, the **palmerpenguins** package has been downloaded over 105,000 times, highlighting the demand and widespread adoption of this viable **iris** alternative. 
 We directly compare the **iris** and **penguins** datasets for selected analyses to demonstrate that R users, in particular teachers and learners currently using **iris**, can switch to the Palmer Archipelago penguins for many use cases including data wrangling, visualization, linear modeling, multivariate analysis (e.g., PCA), cluster analysis and classification (e.g., by k-means).

preamble: |
  % Any extra LaTeX you need in the preamble
output: rticles::rjournal_article
bibiography: packages.bib
---

@@

\bibliography{RJreferences}

>>>>

---
title: Palmer Archipelago Penguins Data in the palmerpenguins R Package - An Alternative to Anderson's Irises
author:
  - name: Allison M. Horst
    affiliation: Bren School of Environmental Science and Management
    address:
    - University of California, Santa Barbara
    - Santa Barbara, CA 93106-5131
    email:  ahorst@ucsb.edu
  - name: Alison Presmanes Hill
    affiliation: RStudio, PBC
    address:
    - 250 Northern Ave
    - Boston, MA 02210
    email:  alison@rstudio.com
  - name: Kristen B. Gorman
    affiliation: College of Fisheries and Ocean Sciences
    address:
    - University of Alaska Fairbanks 
    - 2150 Koyukuk Drive
    - 245 O’Neill Building
    - Fairbanks, AK 99775-7220
    email:  kbgorman@alaska.edu
abstract: > 
 In 1935, Edgar Anderson collected size measurements for 150 flowers from three species of *Iris* on the Gaspé Peninsula in Quebec, Canada. Since then, Anderson's *Iris* observations have become a classic dataset in statistics, machine learning, and data science teaching materials. It is included in the base R **datasets** package as **iris**, making it easy for users to access without knowing much about it. However, the lack of data documentation, presence of non-intuitive variables (e.g. "sepal width"), and perfectly balanced groups with zero missing values make **iris** an inadequate and stale dataset for teaching and learning modern data science skills. Users would benefit from working with a more representative, real-world environmental dataset with a clear link to current scientific research. Importantly, Anderson’s *Iris* dataset appeared in a 1936 publication by R. A. Fisher in the *Annals of Eugenics* (which is often the first-listed citation for the dataset), inextricably linking **iris** to racist research and theory. Thus, a modern alternative to the **iris** dataset is needed. In this paper, we introduce the **palmerpenguins** R package, which includes morphological measurements collected from 2007 - 2009 for three species of *Pygoscelis* penguins that breed on islands throughout the Palmer Archipelago, Antarctica. The **penguins** data provides an approachable, charismatic, and near drop-in replacement for **iris** with topical relevance global climate change and environmental impacts on marine predators. Since the release on CRAN in July 2020, the **palmerpenguins** package has been downloaded over 105,000 times, highlighting the demand and widespread adoption of this viable **iris** alternative. 
 We directly compare the **iris** and **penguins** datasets for selected analyses to demonstrate that R users, in particular teachers and learners currently using **iris**, can switch to the Palmer Archipelago penguins for many use cases including data wrangling, visualization, linear modeling, multivariate analysis (e.g., PCA), cluster analysis and classification (e.g., by k-means). 
preamble: |
  % Any extra LaTeX you need in the preamble
output: rticles::rjournal_article
bibiography: penguins.bib
---

@@

\bibliography{penguins}

### 2021-04-30 penguins.tex
change RJbibliography name to penguins.bib for consistency (#25)
reknit and repurl (#26)

\bibliography{RJreferences}

\address{%
Allison M. Horst\\
Bren School of Environmental Science and Management\\%
University of California, Santa Barbara\\ Santa Barbara, CA 93106-5131\\
%
%
%
\\\href{mailto:ahorst@ucsb.edu}{\nolinkurl{ahorst@ucsb.edu}}
}

\address{%
Alison Presmanes Hill\\
RStudio, PBC\\%
250 Northern Ave\\ Boston, MA 02210\\
%
%
%
\\\href{mailto:alison@rstudio.com}{\nolinkurl{alison@rstudio.com}}
}

\address{%
Kristen B. Gorman\\
College of Fisheries and Ocean Sciences\\%
University of Alaska Fairbanks\\ 2150 Koyukuk Drive\\ 245 O'Neill
Building\\ Fairbanks, AK 99775-7220\\
%
%
%
\\\href{mailto:kbgorman@alaska.edu}{\nolinkurl{kbgorman@alaska.edu}}
}

>>>>

\bibliography{penguins}

\address{%
Allison M. Horst\\
Bren School of Environmental Science and Management\\%
University of California, Santa Barbara\\ Santa Barbara, CA 93106-5131\\
%
%
%
\href{mailto:ahorst@ucsb.edu}{\nolinkurl{ahorst@ucsb.edu}}%
}

\address{%
Alison Presmanes Hill\\
RStudio, PBC\\%
250 Northern Ave\\ Boston, MA 02210\\
%
%
%
\href{mailto:alison@rstudio.com}{\nolinkurl{alison@rstudio.com}}%
}

\address{%
Kristen B. Gorman\\
College of Fisheries and Ocean Sciences\\%
University of Alaska Fairbanks\\ 2150 Koyukuk Drive\\ 245 O'Neill
Building\\ Fairbanks, AK 99775-7220\\
%
%
%
\href{mailto:kbgorman@alaska.edu}{\nolinkurl{kbgorman@alaska.edu}}%
}
